A Data-Driven Framework for Neural Field Modelling

### 2010-11-16 ltx/first_order_framework/LaTeX_Manuscript.tex
changed figures and numbers in the text

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where \dean{is it 50 or 150 with the new experiments?} 150 realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameters converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming that the parameters converged, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

Histograms of the parameter estimates obtained over the \dean{50 or 150} realisations are shown in Figure~\ref{fig:Figure7}. \dean{Can we show the distribution of between 0.85 and 0.95 and put it last in the sequence?} In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, corresponding standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults} with results from Experiment II. \dean{We need to switch the graphs in Figure 10 so that the first one uses these parameters}. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution \dean{(with bias of $2.78\%$) - we need to check this!}, with a small variance in parameter estimates.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for \dean{150} realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(A)), but is missing the higher-frequency detail, which can be see in the error shown in Figure~\ref{fig:Figure10}(C)).

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure6.pdf}
\end{center}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over \dean{150?} realisations. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:Figure6}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure7.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over \dean{150?} realisations}. \dean{David suggested that we switch the order of the graphs here so $\xi$ is the last one.} The solid lines show the actual values of the parameters and the dotted lines show the means of the estimated parameters. (\textbf{A}) The histogram of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. (\textbf{B}) The histogram of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. (\textbf{C}) The histogram of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. (\textbf{D}) The histogram of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Figure7}
\end{figure}

>>>>

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameters converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming that the parameters converged, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, corresponding standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults} with results from Experiment II. \dean{We need to switch the graphs in Figure 10 so that the first one uses these parameters} \parham{it's in ascending order for $\theta_0$ for both rows}. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution (with bias of $2.56\%$), with a small variance in parameter estimates.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(A)), but is missing the higher-frequency detail, which can be see in the error shown in Figure~\ref{fig:Figure10}(C)).

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure6.pdf}
\end{center}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over 150 realisations. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:Figure6}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure7.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over 150 realisations}. The solid lines show the actual values of the parameters and the dotted lines show the means of the estimated parameters. (\textbf{A}) The histogram of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. (\textbf{B}) The histogram of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. (\textbf{C}) The histogram of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$. (\textbf{D}) The histogram of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant.} 
\label{fig:Figure7}
\end{figure}
